---
title: "DSI_slides"
author: "Bowen Liu"
date: "October 7, 2019"
output: html_document
---
#Necessary packages to be installed. 
If the packages have already been installed before, then don't run the following chunk 
```{r}
install.packages("faraway",dependencies = TRUE)
install.packages("ggplot2",dependencies = TRUE) #This might take a while for the first-time R users.  
```

Introduction to data manipulation
1: Read the data (From csv, txt or a specific packages)
For the following example, we're gonna illustrate 3 ways of reading data into R
```{r reading data}
#1. Reading from an existing package (faraway)
library(faraway)
data(gavote)
write.csv(gavote,"gavote.csv",row.names = FALSE)
write.table(gavote,"gavote.txt",row.names = FALSE)
gavote1<-read.csv("gavote.csv")
gavote2<-read.table("gavote.txt")
#Compare gavote1, gavote2 and gavote, they're the same. 
```
# Read the attributes of the data


```{r obtain data info}
gavote
head(gavote)
str(gavote)
summary(gavote)
help(gavote)
#

```

## 

#Plots

```{r Create new variable based on existing variables; plot the histogram and density curves}
gavote$undercount  = (gavote$ballots-gavote$votes)/(gavote$ballots) 
summary(gavote$undercount)
with(gavote,sum(ballots-votes)/sum(ballots))
hist(gavote$undercount,main = "Undercount",xlab = "Percent Undercount")
plot(density(gavote$undercount),main = "Undercount")
rug(gavote$undercount)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r Pie chart and bar plot; equipment of voting}
pie(table(gavote$equip),col = c("pink","red","blue","gray","yellow"))
barplot(sort(table(gavote$equip),decreasing = TRUE),las = 2) #las = 2 indicates the plot is vertical instead of horizontal.
```

#gg plot package and corresponding plots.

```{r ggplot}
gavote$pergore = gavote$gore/gavote$votes
plot(pergore~perAA, gavote,xlab = "proportion of African American", ylab = "proportion for Gore")
plot(undercount~equip,gavote,las = 3)
###### Use ggplot to make the plot look better 
library(ggplot2)
ggplot(gavote) + aes(y = pergore, x = perAA) + geom_point(aes(colour = rural))+geom_smooth(method = lm, se = FALSE, fullrange = TRUE)
ggplot(gavote) + aes(y = undercount, x = equip) +geom_boxplot(outlier.color = c("red"),outlier.shape = c("triangle"),coef = 1.5)
ggplot(gavote) + aes(x = undercount)+geom_histogram(binwidth = 0.001)

```

